---
title: "DESeq analysis "
author: "Valeriia Ladyhina & Semyon Kupriyanov"
date: "`r Sys.Date()`"
output:  
  prettydoc::html_pretty:
    theme: architect
    highlight: lumen
---
# Introduction
*This R Markdown is part of bigger pipeline on analysis of RNA seq data obtained in experiment on effects of different concentrations of antibiotic novobiocin on gene expression in __Salmonella enterica subsp. enterica serovar Typhimurium str. 14028S__. Analysis was performed on publicly available transciptomics data that can be found __[here]('https://www.sciencedirect.com/science/article/pii/S2352340920301918#sec1')__.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Installation of required packages
*This part of pipeline includes few different packages, that have been installed on your computer, namely: __DESeq2, dplyr, ggplot2, gplots__.*
```{r echo=FALSE, warning=FALSE, message=FALSE}
requiredPackages = c('DESeq2', 'dplyr', 'ggplot2','gplots')
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p, repos = "http://cran.us.r-project.org" )
  library(p,character.only = TRUE)}
```
# DESeq2 analysis
*For this analysis we used two files: __transcript_count_matrix.csv__, obtained during previous steps of snalemake pipeline based on StringTie and __pheno.csv__ - file that describes our experimental data.*
```{r echo=FALSE, results='hide', message=FALSE}
countData <- read.csv("./transcript_count_matrix.csv", row.names="transcript_id")
colData <- read.csv('./data/pheno.csv', sep=";", row.names=1)
# Transform to factors
colData$Time <- factor(colData$Time)
colData$Treated <- factor(colData$Treated)
str(colData)
all(rownames(colData) %in% colnames(countData))
countData <- countData[, rownames(colData)]
all(rownames(colData) == colnames(countData))

# Create a DESeqDataSet from count matrix and labels

dds <- DESeqDataSetFromMatrix(countData = countData,
                              colData = colData, design = ~ Treated+Time+Treated:Time)

# Run the default analysis for DESeq2 and generate results table
dds <- DESeq(dds)
res <- results(dds)
# Sort by adjusted p-value and display
(resOrdered <- res[order(res$padj), ])
# Sort the results data frame by the padj and foldChange columns.
sorted = res[with(res, order(padj, -log2FoldChange)), ]
sorted.df = data.frame("id"=rownames(sorted),sorted)
# Write the table out.
write.table(sorted.df, file="result_deseq.txt", sep="\t", col.names=NA, quote=FALSE)
```
```

*Next step is normalization of counts and saving of the result into file __norm-matrix-deseq2.txt__*

```{r echo=FALSE, results='hide'}
# Get normalized counts and write this to a file
# Turn it into a dataframe to have proper column names.
nc = counts(dds,normalized=TRUE)
dt = data.frame("id"=rownames(nc),nc)
# Save the normalize data matrix.
write.table(dt, file="norm-matrix-deseq2.txt", sep="\t",  row.name=FALSE, col.names=TRUE,quote=FALSE)
```

*PCA analysis*

```{r echo=FALSE, results='hide'}
# The function rlogTransform returns a SummarizedExperiment object which contains the rlog-transformed
# values in its assay slot:

rld <- rlog(dds)
head(assay(rld))

plotPCA(rld, intgroup = c("Treated", "Time"))
```

*At this step we are searching for significant genes*

```{r echo=FALSE, results='hide'}
# Find significant genes
res_1 = as.data.frame(res)
res_1 = mutate(res_1, sig=ifelse(res_1$padj<0.05, "FDR<0.05", "Not_sig"))
res_1[which(abs(res_1$log2FoldChange)<1.5), "sig"] = "Not_sig"
res_2 = res_1[which(res_1$sig == "FDR<0.05"),] # 789 genes
sum(res_2$log2FoldChange>0) # 370 positive genes
sum(res_2$log2FoldChange<0) # 419 negative genes
```
*Volcano plot*

```{r echo=FALSE}
# Volcano_plot

ggplot(res_1, aes(log2FoldChange, -log10(padj)))+
  geom_point(aes(col=sig))+
  scale_color_manual(values = c("red", "black"))
```
```{r echo=FALSE, eval=FALSE}
# Check some parametrs
# A so-called MA plot provides a useful overview for an experiment with a two-group comparison:
plotMA(res, ylim = c(-8, 8))
```
```{r echo=FALSE,, eval=FALSE}
# The function plotDispEsts visualizes DESeq2вЂ™s dispersion estimates:
plotDispEsts(dds, ylim = c(1e-6, 1e1))
```
```{r echo=FALSE, eval=FALSE}
# Another useful diagnostic plot is the histogram of the p values
hist(res$pvalue, breaks=20, col="grey")
```

*DEG heatmap*

```{r echo=FALSE}
# Heatmap for DEG
matrix_deg = assay(rld)[rownames(res_2), ]
rownames(matrix_deg) = sapply(rownames(matrix_deg ), 
                    function(x) strsplit(x, split = "_")[[1]][2])
heatmap.2(matrix_deg, scale = "row",
          trace="none", dendrogram = "both")
```

*Negative and positive DEG were saved in files __negative_deg.txt__ and __positive_deg.txt__ respectively.*

```{r echo=FALSE}
# Write negative and positive DEG
write.table(rownames(res_2[res_2$log2FoldChange<0, ]), 
            file="negative_deg.txt", sep="\t", col.names=NA, quote=FALSE)
write.table(rownames(res_2[res_2$log2FoldChange>0, ]), 
            file="positive_deg.txt", sep="\t", col.names=NA, quote=FALSE)
```
