---
title: "KEGG"
author: "Valeriia Ladyhina & Semyon Kupriyanov"
date: "`r Sys.Date()`"
output:  
  prettydoc::html_pretty:
    theme: architect
    highlight: lumen
---
# Introduction
*This R Markdown is part of bigger pipeline on analysis of RNA seq data obtained in experiment on effects of different concentrations of antibiotic novobiocin on gene expression in __Salmonella enterica subsp. enterica serovar Typhimurium str. 14028S__. Analysis was performed on publicly available transciptomics data that can be found __[here]('https://www.sciencedirect.com/science/article/pii/S2352340920301918#sec1')__.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Installation of required packages
*This part of pipeline includes few different packages, that have been installed on your computer, namely: __WGCNA,clusterProfiler, knitr__.*
```{r echo=FALSE, warning=FALSE, message=FALSE}
requiredPackages = c('KEGGREST','stringr','WGCNA','clusterProfiler', 'knitr')
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p, repos = "http://cran.us.r-project.org" )
  library(p,character.only = TRUE)}
```

*In this part we perform KEGG enrichment analysis of modules of interest*

```{r echo=FALSE, message=FALSE, warning=FALSE}
#####Let's make amazing KEGG analysis
yellow = read.table("New_rna_yellow_result.txt")
red = read.table("New_rna_red_result.txt")
black = read.table("New_rna_black_result.txt")
green = read.table("New_rna_green_result.txt")
turquoise = read.table("New_rna_turquoise_result.txt")
brown = read.table("New_rna_brown_result.txt")
blue = read.table("New_rna_blue_result.txt")


kegg_yellow <- enrichKEGG(gene = yellow$V1,
                          organism  = 'seo',
                          pvalueCutoff = 0.05)

kegg_red <- enrichKEGG(gene = red$V1,
                       organism  = 'seo',
                       pvalueCutoff = 0.05)

kegg_black <- enrichKEGG(gene = black$V1,
                         organism  = 'seo',
                         pvalueCutoff = 0.05)

kegg_green <- enrichKEGG(gene = green$V1,
                         organism  = 'seo',
                         pvalueCutoff = 0.05)

kegg_turquoise <- enrichKEGG(gene = turquoise$V1,
                             organism  = 'seo',
                             pvalueCutoff = 0.05)

kegg_blue <- enrichKEGG(gene = blue$V1,
                        organism  = 'seo',
                        pvalueCutoff = 0.05)

kegg_brown <- enrichKEGG(gene = brown$V1,
                         organism  = 'seo',
                         pvalueCutoff = 0.05)
```

*First let's take a look on modules that were sighnificant in relation towards novobiocin treatment*

*Module 1*

```{r echo=FALSE}
kable(head(kegg_turquoise)[2:6])

```

*Module 2*

```{r echo=FALSE}
kable(head(kegg_black)) 
```

*Module 3*

```{r echo=FALSE}
kable(head(kegg_yellow))
```
 
*Module 4*
   
```{r echo=FALSE}
kable(head(kegg_brown))
```
*Module 5*

```{r echo=FALSE}
kable(head(kegg_blue))
```
*And now we will take a look on modules related to time*

*Module 1*

```{r echo=FALSE}
kable(head(kegg_red)[2:6])
```

*Module 2*

```{r echo=FALSE}
kable(head(kegg_green)) 
```

*Now we will perform visualisation of involed KEGG paths and you will find results in your browser*

```{r echo=FALSE}
# Visualization of paths that have been found in both DEGs and modules associated with Treated 
browseKEGG(kegg_black, 'seo02020')
#
browseKEGG(kegg_brown, 'seo00660')
browseKEGG(kegg_brown, 'seo00053')
browseKEGG(kegg_brown, 'seo01210')
#
browseKEGG(kegg_turquoise, 'seo00010')
browseKEGG(kegg_turquoise, 'seo00520')
```
```{r echo=FALSE,message=FALSE,warning=FALSE, results='hide'}

# Download the date
df = read.table("norm-matrix-deseq2.txt", header = T, row.names = 1)
df = as.data.frame(t(df))

# Check the good genes
gsg_df = goodSamplesGenes(df, verbose = 3);
gsg_df$allOK
if (!gsg_df$allOK)

  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg_df$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(df)[!gsg_df$goodGenes], collapse = ", ")));
  if (sum(!gsg_df$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(df)[!gsg_df$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  df = df[gsg_df$goodSamples, gsg_df$goodGenes]

# Make the samples clusterisation
sampleTree = hclust(dist(df), method = "average");

traitData = read.csv('data/pheno.csv', sep=";", row.names=1)
# Samples clustering
sampleTree2 = hclust(dist(df), method = "average")
traitColors = numbers2colors(traitData, signed = F);

powers = c(1:20)
cex1 = 0.8;
# Call the network topology analysis function
sft = pickSoftThreshold(df, powerVector = powers, verbose = 5)

# Build the NET
net = blockwiseModules(df, power = 12,
                       TOMType = "signed", minModuleSize = 30,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "typh_TOM",
                       verbose = 3)

table(net$colors) # 10 moduls + 1 zero

# Visualize
mergedColors = labels2colors(net$colors)


# Some parametrs
moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs;
geneTree = net$dendrograms[[1]];
save(MEs, moduleLabels, moduleColors, geneTree,
     file = "wgcna.RData")

nGenes = ncol(df);
nSamples = nrow(df);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(df, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, traitData, use = "p", method = "spearman");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

# Visual

# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));

# Define variable weight containing the weight column of datTrait
Treated = as.data.frame(traitData$Treated);
names(Treated) = "Treated"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(df, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(df, Treated, use = "p", method = "spearman"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(Treated), sep="");
names(GSPvalue) = paste("p.GS.", names(Treated), sep="");

# OK, Let's define significan modeles (cor>0.7)
# yellow, green, turquoise, black, brown, blue, red

colors_modules = c('yellow', 'green', 'turquoise', 
                   'black', 'brown', 'blue', 'red')
for (i in 1:length(colors_modules)){
  write.table(names(df)[moduleColors==colors_modules[i]], 
              file=paste0(colors_modules[i], "_result.txt"), 
              sep="\t", col.names=NA, quote=FALSE)
}

# Recalculate module eigengenes
MEs = moduleEigengenes(df, moduleColors)$eigengenes
# Isolate weight from the clinical traits
Treated = as.data.frame(traitData$Treated);
names(Treated) = "Treated"
# Add the weight to existing module eigengenes
MET = orderMEs(cbind(MEs, Treated))
```

```{r echo=FALSE}
write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMblack > 0.9),]), 
            file="black_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)

write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMbrown > 0.9),]), 
            file="brown_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)

write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMyellow > 0.9),]), 
            file="yellow_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)

write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMturquoise > 0.9),]), 
            file="turquoise_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)


# Time hubs
write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMgreen > 0.9),]), 
            file="green_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)

write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMred > 0.9),]), 
            file="red_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)

write.table(rownames(geneModuleMembership[which(geneModuleMembership$MMblue > 0.9),]), 
            file="blue_hub.txt", 
            sep="\t", col.names=NA, quote=FALSE)
```

