---
title: "KEGG and WGCNA"
author: "Valeriia Ladyhina & Semyon Kupriyanov"
date: "`r Sys.Date()`"
output:  
  prettydoc::html_pretty:
    theme: architect
    highlight: lumen
---
# Introduction
*This R Markdown is part of bigger pipeline on analysis of RNA seq data obtained in experiment on effects of different concentrations of antibiotic novobiocin on gene expression in __Salmonella enterica subsp. enterica serovar Typhimurium str. 14028S__. Analysis was performed on publicly available transciptomics data that can be found __[here]('https://www.sciencedirect.com/science/article/pii/S2352340920301918#sec1')__.*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Installation of required packages
*This part of pipeline includes few different packages, that have been installed on your computer, namely: __clusterProfiler, WGCNA, knitr__.*
```{r echo=FALSE, warning=FALSE, message=FALSE}
requiredPackages = c('clusterProfiler','WGCNA','knitr')
for(p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p, repos = "http://cran.us.r-project.org" )
  library(p,character.only = TRUE)}
set.seed(7)
```
# KEGG enrichment analysis
*Negative DEG*
```{r echo=FALSE, results='hold', message=FALSE, warning=FALSE}
# Download transformed deg-lists
deg_negative = read.table("New_rna_negative_deg.txt")
deg_positive = read.table("New_rna_positive_deg.txt")

# KEGG enrichmant analysis
kegg_deg_negative <- enrichKEGG(gene = deg_negative$V1,
                         organism  = 'seo',
                         pvalueCutoff = 0.05)

kegg_deg_positive <- enrichKEGG(gene = deg_positive$V1,
                         organism  = 'seo',
                         pvalueCutoff = 0.05)
# Search results
kable(head(kegg_deg_negative)) # Flagellar assembly # Oxidative phosphorylation
# Glycolysis / Gluconeogenesis # Bacterial chemotaxis
# Starch and sucrose metabolism

```

*Positive DEG*
```{r echo=FALSE}
kable(head(kegg_deg_positive)) 
```


```{r echo=FALSE}
# Download the date
df = read.table("norm-matrix-deseq2.txt", header = T, row.names = 1)
df = as.data.frame(t(df))
```
```{r echo=FALSE, results='hide'}
# Check the good genes
gsg_df = goodSamplesGenes(df, verbose = 3);
gsg_df$allOK
if (!gsg_df$allOK)
{
  # Optionally, print the gene and sample names that were removed:
  if (sum(!gsg_df$goodGenes)>0)
    printFlush(paste("Removing genes:", paste(names(df)[!gsg_df$goodGenes], collapse = ", ")));
  if (sum(!gsg_df$goodSamples)>0)
    printFlush(paste("Removing samples:", paste(rownames(df)[!gsg_df$goodSamples], collapse = ", ")));
  # Remove the offending genes and samples from the data:
  df = df[gsg_df$goodSamples, gsg_df$goodGenes]
}
```

# Sample clusterisation to detect outriers

```{r echo=FALSE}
# Make the samples clusterisation
sampleTree = hclust(dist(df), method = "average");
# Plot the sample tree: Open a graphic output window of size 12 by 9 inches
plot(sampleTree, main = "Sample clustering to detect outliers", sub="", xlab="", cex.lab = 1.5,
     cex.axis = 1.5, cex.main = 2)
```

*Added to clusters information regarding traits of data set*

```{r echo=FALSE}
# Download traits
traitData = read.csv('data/pheno.csv', sep=";", row.names=1)
# Samples clustering
sampleTree2 = hclust(dist(df), method = "average")
traitColors = numbers2colors(traitData, signed = F);
plotDendroAndColors(sampleTree, traitData,
                    groupLabels = names(traitData),
                    main = "Sample dendrogram and trait heatmap")
```
```{r echo=FALSE, results='hide', warning=FALSE}
# Pick soft threshold
powers = c(1:20)
cex1 = 0.8;
# Call the network topology analysis function
sft = pickSoftThreshold(df, powerVector = powers, verbose = 5)
```
*In this case we will use soft threashold in power 12*
```{r echo=FALSE}
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,cex=cex1,col="red");
# this line corresponds to using an R^2 cut-off of h
abline(h=0.80,col="red")
# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
     xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
     main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")
```
```{r echo=FALSE, results='hide'}
# Build the NET
net = blockwiseModules(df, power = 12,
                       TOMType = "signed", minModuleSize = 30,
                       reassignThreshold = 0, mergeCutHeight = 0.25,
                       numericLabels = TRUE, pamRespectsDendro = FALSE,
                       saveTOMs = TRUE,
                       saveTOMFileBase = "typh_TOM",
                       verbose = 3)

table(net$colors) # 10 moduls + 1 zero
```

*Module clustering dendrogram* 

```{r echo=FALSE}
# Visualize
mergedColors = labels2colors(net$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(net$dendrograms[[1]], mergedColors[net$blockGenes[[1]]],
                    "Module colors",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = TRUE, guideHang = 0.05)


# Some parametrs
moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
MEs = net$MEs;
geneTree = net$dendrograms[[1]];
save(MEs, moduleLabels, moduleColors, geneTree,
     file = "wgcna.RData")
```
*Significance of modules and their correlation with traits using Spearman's correlation, we will count that resulta are significant if the are >0.7*

```{r echo=FALSE, warning=FALSE}
# Define modules significance and correlation with Traits

# Define numbers of genes and samples
nGenes = ncol(df);
nSamples = nrow(df);
# Recalculate MEs with color labels
MEs0 = moduleEigengenes(df, moduleColors)$eigengenes
MEs = orderMEs(MEs0)
moduleTraitCor = cor(MEs, traitData, use = "p", method = "spearman");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

# Visual

# Will display correlations and their p-values
textMatrix = paste(signif(moduleTraitCor, 2), "\n(",
                   signif(moduleTraitPvalue, 1), ")", sep = "");
dim(textMatrix) = dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3));
# Display the correlation values within a heatmap plot
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(traitData),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = greenWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1,1),
               main = paste("Module-trait relationships"))
```
*In our analysis we have five modules that are significant regarding treatment: brown, turquoise, blue, black, yellow; and two regarding time green and red*

*Plot that reveals relationships between eigengenes and traits*

```{r echo=FALSE}
# Define variable weight containing the weight column of datTrait
Treated = as.data.frame(traitData$Treated);
names(Treated) = "Treated"
# names (colors) of the modules
modNames = substring(names(MEs), 3)
geneModuleMembership = as.data.frame(cor(df, MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));
names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");
geneTraitSignificance = as.data.frame(cor(df, Treated, use = "p", method = "spearman"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
names(geneTraitSignificance) = paste("GS.", names(Treated), sep="");
names(GSPvalue) = paste("p.GS.", names(Treated), sep="");

# OK, Let's define significan modeles (cor>0.7)
# yellow, green, turquoise, black, brown, blue, red

colors_modules = c('yellow', 'green', 'turquoise', 
                   'black', 'brown', 'blue', 'red')
for (i in 1:length(colors_modules)){
  write.table(names(df)[moduleColors==colors_modules[i]], 
              file=paste0(colors_modules[i], "_result.txt"), 
              sep="\t", col.names=NA, quote=FALSE)
}

# Recalculate module eigengenes
MEs = moduleEigengenes(df, moduleColors)$eigengenes
# Isolate weight from the clinical traits
Treated = as.data.frame(traitData$Treated);
names(Treated) = "Treated"
# Add the weight to existing module eigengenes
MET = orderMEs(cbind(MEs, Treated))
# Plot the relationships among the eigengenes and the trait
par(cex = 0.9)
plotEigengeneNetworks(MET, "", marDendro = c(0,4,1,2), marHeatmap = c(3,4,1,2), cex.lab = 0.8, xLabelsAngle
                      = 90)
```